---
title: "Week 1 Notes"
author: "Erik Rehnberg Steeb"
date: "3/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Selection

Take regression - select highest R^2 - equivalent to selecting model with lowest Mean Squared Error.

Overfitting danger - as variables are added, we can be fitting observed data - gives high R^2, low MSE, awful forecasts - not accurately modeling trends. Adding parameters reduces degrees of freedom.

Also looking for lowest s^2. s^2 as penalty weight (T/(T-k))MSE - new parameter will lower MSE, but will increase (T/(T-k)) - s^2 may or may not decrease.

```{r cars}
summary(cars)
```

## Lag Operator
* Accounts for how changes may not immediately affect measures. For example, tax policy may not immediately impact GDP, since people will react over time, not instantaneously.
** Need to determine how far back things have effects - convention is Lyt = yt-1 (some books use B). Two period lag --> L^2y = yt-2
