---
title: "Project Setup"
author: "Erik Rehnberg Steeb"
date: "3/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)

# Change this line to reflect your path/comment it out if you don't need to change your wd
setwd("./Project")
```

## Read in files

Read in 2021 data (so far). Each worksbook has 3 tabs/worksheets. Production is tab/sheet 2, curtailment is sheet 3. There's a generic readme on sheet 1, which we can ignore

```{r}
production_data <- read_excel("data/ProductionAndCurtailmentsData_2021.xlsx",sheet = 2)
head(production_data)
```

This next section will read in all workbooks and output two dataframes, `{r} production_df` and `{r}curtailment_df`. The `Load` through `Thermal` columns of the production tabs each contain commas in the columns, which need to be cleaned. Similarly, the curtailment column reads in a host of `NAs`, as well as containing commas. 

```{r}
files <- list.files("data/")

production_df <- data.frame()
curtailment_df <- data.frame()

for (i in 1:length(files)){
  prod_data <- read_excel(paste0("data/",files[i]), sheet = 2)
  curt_data <- read_excel(paste0("data/",files[i]), sheet = 3)
  production_df <- rbind(production_df, prod_data)
  curtailment_df <- rbind(curtailment_df, curt_data)
  print(i)
  i <- i+1
}


```